\documentclass[overlapped,line,letterpaper]{res}

\usepackage{ifpdf}

\ifpdf
  \usepackage[pdftex]{hyperref}
\else
  \usepackage[hypertex]{hyperref}
\fi

\hypersetup{
  letterpaper,
  colorlinks,
  urlcolor=black,
  pdfpagemode=none,
  pdftitle={Curriculum Vitae},
  pdfauthor={Damjan Škulj},
  pdfsubject={Curriculum Vitae},
}

%%===========================================================================%%

\begin{document}
	
	%---------------------------------------------------------------------------
	% Document Specific Customizations
	
	% Make lists without bullets and with no indentation
	\setlength{\leftmargini}{0em}
	\renewcommand{\labelitemi}{}
	
	% Use large bold font for printed name at top of pages
	\renewcommand{\namefont}{\large\textbf}
	
	%---------------------------------------------------------------------------
	
	\name{Damjan \v{S}kulj}
	
	
	
	\begin{resume}
		Damjan \v{S}kulj was born at 11 March 1975 in Novo Mesto, Slovenia. He studied theoretical mathematics at the Faculty of Mathematics and Physics of University of Ljubljana and graduated in 1998. At the same faculty he also finished his Master degree in 2001 and obtained PhD in 2004. In 1999 he attended the Essex Summer School in Social Science Data Analysis at University of Essex, UK. 
		
		He worked as a teaching assistant at the Faculty of Social Sciences of University of Ljubljana (FDV, UL) between 1998 and 2006 and as an assistant professor at the same institution between 2006 and 2016. In 2016 he was promoted to the degree of associate professor. He taught subjects \textit{Mathematics, Statistics, Operational research}. He was a teaching assistant for {\em Maths refresher course} at the ECPR Summer School in Methods \& Techniques, University of Ljubljana, 2007 and a teacher for {\em The most essential topics in Probability Theory} at the ECPR Summer School in Methods \& Techniques, University of Ljubljana, 2010. In 2010 he also taught the course {\em Imprecise Markov Chains} at the 4th SIPTA Summer School in Durham, UK. 
		
		In 2008 he received the Alan Richards Fellowship at Grey College, Durham University, UK, April to June 2008.
		
		He was a head of the Chair of Social informatics and Methodology at FDV, UL 2006--2008 and 2011--2013. 
		
		He is a member of the Society of Mathematicians, Physicists and Astronomers of Slovenia; the Slovenian Society Informatika; the Statistical Society of Slovenia and the Society for Imprecise Probability: Theories and Applications. 
		
		He is a reviewer for Mathematical reviews, a member of editorial board at International Journal of Data Analysis Techniques and Strategies and a reviewer for several journals and scientific conferences.
		
		His research interests are mainly related to imprecise probabilities. They include modelling stochastic processes under uncertainty, both discrete and continuous; modelling risk and uncertainty for decision making; differential equations with uncertain parameters; computational aspects of stochastic models.  
		
		
		
		
		
			
		\section{\bf Selected publications}
		
		``Discrete time Markov chains with interval probabilities.'' {\em International Journal of Approximate Reasoning}, 2009.
		
		``Coefficients of ergodicity for Markov chains with uncertain parameters.'' with R. Hable  {\em Metrika}, 2013.
		
		``A classification of invariant distributions and convergence of imprecise Markov chains.''  {\em  Linear Algebra and its Applications}, 2013.
		
		``Efficient computation of the bounds of continuous time imprecise Markov chains.'' {\em  Applied Mathematics and Computation}, 2015. 
		
		``Random walks on graphs with interval weights and precise marginals.'' {\em International Journal of Approximate Reasoning}, 2016.		
	\end{resume}
	
\end{document}